\Title

### 2022-09-27 Grimaldi-etal-BiolCybernetics.tex
Introduction: 1.1

%\ninept
%
%\abstract{Abstract}
%
\keywords{\Keywords}
%
\maketitle
%%%%%%%%% BODY TEXT
\section{Introduction}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
\label{sec:intro}
% >>>>>>>>>>> LuP is here <<<<<<<<<<<
%
\subsection{Recent challenges in machine learning and computer vision}
%%%-----------------------------------------------------------------
Machine learning, and computer vision in particular, have undergone a major transformation in modern times, leaving science fiction movies to become mainstream tools found in the smartphones in everyone's pocket. Less visible, but even more crucial to our economy and security, are the many automated tools that now dominate the range of services that equip virtually all sectors of our daily lives. This trend is most pronounced in the field of medicine, as it provides practitioners with new diagnostic tools to treat people faster and more effectively. This particular example also illustrates some important challenges that these tools are increasingly facing. One of these is their sensitivity to adversarial attacks, that is, the sensitivity of the models' output to minute changes in the input. This could raise important questions about the reliability of the results of in any automated diagnostic model. More generally, these models do not explain why they produce a given output for a class of inputs. This lack of explainability is a hindrance to the advancement of these methodologies as well as to their acceptance for most safety-relevant tasks.

If we take a step back, we will remember that these tools are inherited from theoretical models designed to understand the brain. While the first artificial neural networks were mainly applied to " toy " type problems, they also offered a growing number of applications in computer vision, but also more generally for task automation. In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates, and today they reach higher performances than humans for image categorization. One of the challenges of these models is to be able to scale them up for future applications, for example in autonomous driving. Indeed, in such applications, the input does not consist of a single image (say a few million pixels) but of a continuous stream of images from several cameras.

@@

Second, a major constraint on future systems is energy budget. As we saw, a fast computation time was maintained for high-dimension images, but this come at the cost of a higher energy consumption. Also, the need to solve increasingly more complex tasks necessitated the construction of increasingly more complex and deeper architectures. As en example, the top performing model on the Imagenet challenge is using billions of parameters which typically cost 1Wh to train and approximately 1Wh per inference of a single image. Less greedy solutions have emerged such as Mobile Net, yet these solutions are yet not ready to be used in practice in autonomous cars, as the added consumption would be approximately 300W that is, 20\% of the energy of an electric can in standard conditions.

Neuroscience could bring breakthrough answers to these two intertwined problems. Indeed, the human brain has the remarkable property of being able to act and respond at any time, while consuming a reasonable amount of energy, estimated to 5W at rest. This system is the result of million of years of natural selection, and a striking difference of biological neural networks compared to an artificial one is the representation that they use. CNNs for instance represent the information passing from one layer to the other as tensors representing visual information in a dense manner on the visual topography, while representing different properties in different channels. On the other side, information in a vast majority of biological neural networks is represented as spikes, that is, as prototypical all-or-none events whose only parameters are their timing and the address of the neuron that emitted that spite. Why spikes ? It is not yet known the reason why most systems evolved to use that representation, but one major advantage seems to be a form of efficiency. In the human brain, each of the about ten trillion nerve cells emit on average 1 spike per second, but each neurons fires with a high variability, and up to 300 spikes per second. There has been numerous attempts to build artificial spiking neural networks, and notably to have the same performance as that observed in classical CNNs, yet with a lower energy budget. Yet, no such SNN has yet surpassed state-of the art performances. Here, we will attempt to focus on a neurobiologically inspired property of spiking neurons to advance in that endeavor.

\subsection{What can neuroscience bring to computer vision?}% and the Hetero-synaptic delays in spiking neurons}
%%%-----------------------------------------------------------------
% 1.2.a
Let's focus first on some existing spiking neural network models. one first motivation for these models was to understand the biological observations at different scales. For instance the Hodgkin Huxley mode describes the 'dynamic of the membrane potential and the role of the different voltage-gated ion channels. It also explains how the all may emit a spike. Larger scale models introduce an explicit spiking mechanism, for instance in the Integrate-and-Fine model which allows to predict the emerging states of a network of recurrently connected neurons (Brunel). Several models have sealed up the system to include more neurons such as the Virtual Brain (Sanz) which allows to mix spiking neural nets with analog neural mass models. However these methods are merely descriptive and have application in neuroscience and few have had an impact in computer vision.

>>>>

%\ninept
%
%\abstract{Abstract}
%
\keywords{\Keywords}
%
\maketitle
%%%%%%%%% BODY TEXT
\section{Introduction}
%%%-----------------------------------------------------------------
%%%-----------------------------------------------------------------
\label{sec:intro}
% >>>>>>>>>>> LuP is here <<<<<<<<<<<
%
\subsection{Recent challenges in machine learning and computer vision}
%%%-----------------------------------------------------------------
Machine learning, and computer vision in particular, have undergone a major transformation in modern times, leaving science fiction movies to become mainstream tools found in the smartphones in everyone's pocket. Less visible, but even more crucial to our economy and security, are the many automated tools that now dominate the range of services that equip virtually all sectors of our daily lives. This trend is most pronounced in the field of medicine, as it provides practitioners with new diagnostic tools to treat people faster and more effectively. This particular example also illustrates some important challenges that these tools are increasingly facing. One of these is their sensitivity to adversarial attacks, that is, the sensitivity of the models' output to minute changes in the input. This could raise important questions about the reliability of the results of in any automated diagnostic model. More generally, these models do not explain why they produce a given output for a class of inputs. This lack of explainability is a hindrance to the advancement of these methodologies as well as to their acceptance for most safety-relevant tasks. If we take a step back, we will remember that these tools are inherited from theoretical models designed to understand the brain. While the first artificial neural networks were mainly applied to " toy " type problems, they also offered a growing number of applications in computer vision, but also more generally for task automation. In particular, convolutional neural networks are inspired by the feed-forward processing observed in the visual pathways of primates, and today they reach higher performances than humans for image categorization. One of the challenges of these models is to be able to scale them up for future applications, for example in autonomous driving. Indeed, in such applications, the input does not consist of a single image (say a few million pixels) but of a continuous stream of images from several cameras.

@@

Second, a major constraint for future systems is energy cost. As we have seen, fast computation time has been maintained for high-dimensional images, but at the cost of higher energy consumption. In addition, the need to solve increasingly complex tasks has required the construction of increasingly complex and deep architectures. As an example, the best performing model in the Imagenet challenge uses billions of parameters that typically cost 1Wh to train and about 1Wh to infer a single image. Less power-hungry solutions have emerged, such as Mobile Net, but they are not yet ready for practical use in autonomous cars, as the additional power consumption would be about 300W, or 20\% of the energy of an electric car under standard conditions.
Neuroscience could provide revolutionary answers to these two closely related problems. Indeed, the human brain has the remarkable property of being able to react at any time, while consuming a reasonable amount of energy, estimated at 5W at rest. This system is the result of millions of years of natural selection, and a striking difference between biological neural networks and artificial ones is the representation they use. CNNs, for example, represent information passing from one layer to another as tensors representing visual information densely over the visual topography, while representing different properties in different channels. On the other hand, information in a large majority of biological neural networks is represented as spikes, i.e. prototypical all-or-nothing events whose only parameters are their timing and the address of the neuron that emitted this spike. Why spikes? It is not yet known why most biological systems have evolved to use this representation, but one major advantage seems to be some form of efficiency. In the human brain, each of the approximately ten trillion nerve cells emits on average one spike per second, but each neuron fires with great variability, and up to 300 spikes per second. Many attempts have been made to build artificial spiking neural networks, and in particular to achieve the same performance as observed in classical CNNs, but with a lower energy budget. However, none of these networks has yet surpassed the state of the art performances. Here, we will try to focus on a property of spike neurons inspired by neurobiology to progress in this direction.
\subsection{What can neuroscience bring to computer vision?}% and the Hetero-synaptic delays in spiking neurons}
% 1.2.a
Let's focus first on some existing spiking neural network models. one first motivation for these models was to understand the biological observations at different scales. For instance the Hodgkin Huxley mode describes the 'dynamic of the membrane potential and the role of the different voltage-gated ion channels. It also explains how the all may emit a spike. Larger scale models introduce an explicit spiking mechanism, for instance in the Integrate-and-Fine model which allows to predict the emerging states of a network of recurrently connected neurons (Brunel). Several models have sealed up the system to include more neurons such as the Virtual Brain (Sanz) which allows to mix spiking neural nets with analog neural mass models. However these methods are merely descriptive and have application in neuroscience and few have had an impact in computer vision.
